---
Title: Comparison electrode placement
Author: Filip Niemann
Date: 22.04.2024
output: html_document
---


# This script compares data for errors in  neuronavigated electrode 
# placement with (citation JoVE) and non neuronavigated electrode 
# placement doi: https://doi.org/10.1101/2024.02.19.24302917
# for details of the datasets please view the original paper.
# remind, data can be different then in the original JoVE paper since
# it was an ongoing project and data tables were updated to newest versions.
# Type of analysis didn't changed.



```{r, eval = requireNamespace("tidyverse", quietly = TRUE)}
library(here)
library(ggeffects)
library(dplyr)
library(ggpubr)
library(lme4)
library(car)
library(patchwork)
library(gridExtra)
library(cowplot)
library(tidyverse)
library(grid)
library(lattice)
library(ggplot2)
library(arrow)
library(dplyr)
library(rstatix)
library(emmeans)
library(knitr)
library(lme4)
library(afex)
library(tidyr)
library(sjmisc)
library(lmerTest)
library(httpgd)
library(r2glmm)
library(see)

```

## Code style
### read in for LMM
```{r}
path <- here()
add <- "06_Comparison_Electrode_Placement_Error"
df <- read.csv(file.path(path, add, "df_dim_norm_diff.csv"))
df <- df %>% mutate(across("run", str_replace, 'baseline', 'intended'))
df <- df %>% mutate(across("run", str_replace, 'pre fMRI', 'actual pre-fMRI'))
df <- df %>% mutate(across("run", str_replace, 'post fMRI', 'actual post-fMRI'))
df <- df[!grepl('sub-002', df$Subject), ]
View(df)
```


```{r}
keeps <- c("Subject", "Region", "Session", "run", "Diff_Norm")
df <- df[keeps]
df <- df[!is.na(df$Diff_Norm), ]
View(df)
```

```{r}
ggboxplot(df, x = "run", y = "Diff_Norm", color = "Session")

```

## identify outliers
```{r}
df %>% group_by(run, Region) %>% identify_outliers(Diff_Norm)
```

# Relevant code
```{r}
df %>%
    group_by(Region) %>%
    summarise_all(count, na.rm = TRUE) %>%
    filter(run == 'intended')
```

```{r}
df_test <- df %>% filter(run != 'intended')
View(df_test)
lower_bound <- quantile(df[, 'Diff_Norm'], 0.025)
upper_bound <- quantile(df[, 'Diff_Norm'], 0.975)
outlier_ind <- which(df$Diff_Norm < lower_bound | df$Diff_Norm > upper_bound)
valid_sub <- which(df$Diff_Norm > lower_bound | df$Diff_Norm < upper_bound)
outlier_ind
```

```{r}
x <- quantile(df[, 'Diff_Norm'], c(0.025, 0.975))
df <- df[df$Diff_Norm >= x[1] & df$Diff_Norm <= x[2], ]
View(df)
```

```{r}
df[outlier_ind, ]
```

```{r}
View(df[valid_sub, ])
```

```{r}
df <- df[valid_sub]
```

```{r}
a <- length(unique(df$Subject))
print(a)
```

## get age and sex
```{r}
list_sub <- unique(df$Subject)

df_part <- read.csv(file.path(path, add, "participant.tsv"), sep = "\t") # original
df_subset <- subset(df_part, participant_id %in% list_sub)
```

```{r}
length(unique(df_subset$participant_id))
length(which(df_subset$sex == 'male'))
length(which(df_subset$sex == 'female'))
length(which(df_subset$sex == 'non binary'))

```

## count number of datapoints
```{r}
a <- length(unique(df$Subject))
print(a)

count(df %>% filter(run == 'planned post-fMRI'))
count(df %>% filter(run == 'planned pre-fMRI'))
count(df %>% filter(run == 'intended'))
```
```{r}
count(df %>% filter(run == 'planned post-fMRI', Region == 'rOTC'))
count(df %>% filter(run == 'planned pre-fMRI', Region == 'rOTC'))  #"rOTC", "rDLPFC", "lTPC"
```

```{r}
count(df %>% filter(run == 'planned post-fMRI', Region == 'rDLPFC'))
count(df %>% filter(run == 'planned pre-fMRI', Region == 'rDLPFC'))
```

```{r}
count(df %>% filter(run == 'planned pre-fMRI', Region == 'lTPC'))
count(df %>% filter(run == 'planned post-fMRI', Region == 'lTPC'))
```


```{r}
df %>%
  group_by(Region) %>%
  summarise(count = n_distinct(Subject))

```

```{r}
df %>%
  group_by(Region) %>%
  summarise(count = n_distinct(Session))

```

## Normality assumptions

```{r}
df %>% group_by(run, Region) %>% shapiro_test(Diff_Norm)
```

```{r}
ggqqplot(df, "Diff_Norm", ggtheme = theme_bw()) + facet_grid(Session ~ Region)
```

## Homogneity of variance assumption
```{r}
df %>% group_by(Region) %>% levene_test(Diff_Norm ~ run)
```

# choose factor, first factor values dummy coded as 0

```{r}
df$Experiment <- factor(df$Experiment, levels = c("rOTC", "rDLPFC", "lTPC"))
df$Session <- factor(df$Session, levels = c("ses-0", "ses-1", "ses-2", "ses-3", "ses-4"))
df$run <- factor(df$run, levels = c("intended", "actual pre-fMRI", "actual post-fMRI"))
```

# create mixed models
```{r}
# help define within and between factor
# https://www.frontiersin.org/journals/psychology/articles/10.3389/fpsyg.2015.00002/full
lmer_model_1 <- lmer(Diff_Norm ~ 1 + run * Region + (1 | Subject), data = df)
lmer_model_2 <- lmer(Diff_Norm ~ 1 + run * Region + (1 + Region| Subject), data = df)
lmer_model_3 <- lmer(Diff_Norm ~ 1 + run * Region + (1 + run | Subject), data = df)
```

## model comparison
```{r}
anova(lmer_model_1,lmer_model_2,lmer_model_3)
```

## plot table of winning model

```{r}
sjPlot::tab_model(lmer_model_1, show.ci = 0.95, show.p = TRUE, show.df = TRUE)
sjPlot::tab_model(lmer_model_1, show.ci = 0.95, show.p = TRUE, show.df = TRUE, file = file.path(path, add, "tables","model5.docx"))

```

## Calculate estimate marginal means (EMMs)

### main effect Region
```{r}
# Anova of winning model
(EMM.Region <- emmeans(lmer_model_1, "Region"))
```
### main effect position (run)
```{r}
# Anova of winning model
(EMM.run <- emmeans(lmer_model_1, "run"))
```

### simple effect region for each level of position (run)

```{r}
emmeans(lmer_model_1, pairwise ~ Region|run)
```

### simple effect position (run) for each level of region
```{r}
emmeans(lmer_model_1, pairwise ~ run|Region)
```

```{r}
emmip(lmer_model_1, ~ run | Region, CIs = TRUE, type='response') +
    geom_point(aes(x = run, y = Diff_Norm), data = df, pch = 1, color = "blue", size=1.5)
```

```{r}
emmip(lmer_model_1, ~ run, CIs = TRUE) +
    geom_point(aes(x = run, y = Diff_Norm), data = df, pch = 1, color = "#5b5ba1", size=5.5) + 
    theme(text = element_text(size = 45),
    axis.text.x = element_text(angle = 0, hjust = 1)) + 
    labs(y= "EEMs L² Norm (intended - actual) in |mm|", x = "Time point")

```

```{r}
emmeans(lmer_model_1, pairwise ~ run)
```

```{r}
emmip(lmer_model_1, ~ Region, CIs = TRUE) +
    geom_point(aes(x = Region, y = Diff_Norm), data = df, pch = 1, color = "#5b5ba1", size=5.5) + 
    theme(text = element_text(size = 45),
    axis.text.x = element_text(angle = 0, hjust = 1)) + 
    labs(y = "EEMs L² Norm (intended - actual) in |mm|", x = "Time point")

```

```{r}
emmeans(lmer_model_1, pairwise ~ Region)
```

```{r}
emmip(lmer_model_1, ~ run | Region, CIs = TRUE) +
    geom_point(aes(x = run, y = Diff_Norm), data = df, pch = 1, color = "#0707ec", size=5.5) + 
    theme(text = element_text(size = 45),
    axis.text.x = element_text(angle = 45, hjust = 1)) + 
    labs(y = "EEMs L² Norm (intended - actual) in |mm|", x = "Position")
```



```{r}
p1 <- emmip(lmer_model_1, ~ run, CIs = TRUE) +
    geom_point(aes(x = run, y = Diff_Norm), data = df, pch = 1, color = "#15b7e9", size=1.5) + 
    theme(text = element_text(size = 22),
        axis.text.x = element_text(angle = 0)) + 
        labs(y = "EMMs L² Norm (intended - actual) in |mm|", x = "Position")
png(file.path(path, add, "figures", "Figure_6_R.png"))
print(p1)
dev.off()  
```

```{r}
p2 <- emmip(lmer_model_1, ~ Region | run, CIs = TRUE) +
    geom_point(aes(x = Region, y = Diff_Norm), data = df, pch = 1, color = "#15b7e9", size=1.5) + 
    theme(text = element_text(size = 22),
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
        labs(y = "EMMs L² Norm (intended - actual) in |mm|", x = "Region")
png(file.path(path,add, "figures", "Figure_7_R.png"))
print(p2)
dev.off()         
```

```{r}
p3 <- emmip(lmer_model_1, ~  run | Region, CIs = TRUE) +
    geom_point(aes(x = run, y = Diff_Norm), data = df, pch = 1, color = "#15b7e9", size=1.5) + 
    theme(text = element_text(size = 22),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
        labs(y = "EMMs L² Norm (intended - actual) in |mm|", x = "Position")
png(file.path(path, add, "figures", "Figure_8_R.png"))
print(p3)
dev.off()         
```

```{r}
p4 <- emmip(lmer_model_1, ~  Region, CIs = TRUE) +
    geom_point(aes(x = Region, y = Diff_Norm), data = df, pch = 1, color = "#15b7e9", size=1.5) + 
    theme(text = element_text(size = 22),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
        labs(y = "EMMs L² Norm (intended - actual) in |mm|", x = "Position")
png(file.path(path, add, "figures", "Figure_9_R.png"))
print(p4)
dev.off()
```