---
title: "R script for SimNIBS simulated normal component E-field values "
author: Filip Niemann
date: 20.09.2023
output: html_document
---

```{r}
library(performance)
library(lattice)
library(ggplot2)
library(arrow)
library(dplyr)
library(rstatix)
library(emmeans)
library(knitr)
library(lme4)
library(afex)
library(tidyr)
library(sjmisc)
library(lmerTest)
library(httpgd)
library(r2glmm)
library(see)
library(hrbrthemes)
library(ggpubr)    # For combining graphs but install not workin
theme_set(theme_pubr())
library(patchwork)
library(gridExtra)
library(cowplot)
library(grid)
# If you use visual code, images will be shown in extra browser window
hgd()
hgd_browse()
```

```{r, echo=FALSE}
library(rstudioapi)
source_path <- dirname(getSourceEditorContext()$path)
# Should be the path on your local machine
View(source_path)
# Use given functions to calculate estimated marginal means and plot them
source(file.path(source_path, "plot_emmip.R"))
# Path to figure and table files (some are given, some will be created)
path_source_data <- (file.path(source_path, "Source_Data"))
```

```{r}
# Style and path variables.
options(digits = 5)
digits <- 2
digits2 <- 4
path_images <- file.path(source_path, "Images_and_Tables")
width <- 8
height <- 7
faktor <- 0.7
dpi <- 150
fontsize <- 28
fontsize_2 <- 26
ylab_text <- "mean nE in V/m"
ylab_LMM_text <- "95% CI EMM in V/m"
```

# 1. Data preparation
## Read in tables
- Long format for simulated normal component of positiv E-field values of actual and planned Electrodposition for each Montage and Area

```{r}
#Normal component positive part meanE

# Normal components 
M1_readout_M1_normal_meanE_125 <- "/df_sphere_E_normal_fsaverage_mni_space_12.5_IFG_origin_IFG_M1_origin_M1_readoutrM1.feather"
M1_readout_M1_normal_meanE_25 <- "/df_sphere_E_normal_fsaverage_mni_space_25_IFG_origin_IFG_M1_origin_M1_readoutrM1.feather"
# (Optional)
M1_readout_M1_normal_meanE_375 <- "/df_sphere_E_normal_fsaverage_mni_space_37.5_IFG_origin_IFG_M1_origin_M1_readoutrM1.feather"


data_meanE_normal_125 <- paste0(path_source_data, M1_readout_M1_normal_meanE_125)
data_meanE_normal_25 <- paste0(path_source_data, M1_readout_M1_normal_meanE_25)
# (Optional)
data_meanE_normal_375 <- paste0(path_source_data, M1_readout_M1_normal_meanE_375)

df_meanE_normal_125 <- arrow::read_feather(data_meanE_normal_125)
df_meanE_normal_25 <- arrow::read_feather(data_meanE_normal_25)
# (Optional)
df_meanE_normal_375 <- arrow::read_feather(data_meanE_normal_375)
```
# create Area column from Experiment column
```{r}
View(df_meanE_normal_25)

df_meanE_normal_125$sphere <- "1.25"
df_meanE_normal_25$sphere <- "2.5"
# (Optional)
df_meanE_normal_375$sphere <- "3.75"
```


```{r}
df <- rbind(df_meanE_normal_125, df_meanE_normal_25)

df <- df %>% mutate(Area = case_when(
    grepl(
    pattern = "IFG", x = Experiment) ~ "IFG",
    grepl(
    pattern = "M1", x = Experiment) ~ "M1")
    )
```
```{r}
# Delete additional not used experimental data
df <- df[!grepl("IFG \n conv Phon", df$Experiment), ]
View(df)
```

# Change Montage

```{r}
df <- within(df, rm(Montage))
df <- df %>% mutate(Montage = case_when(
    grepl(
    pattern = "conv", x = Experiment) ~ "conv",
    grepl(
    pattern = "focal", x = Experiment) ~ "focal")
    )
```



# create index for random variable
```{r}
list_nested <- c("Experiment", "id")
lmm_group <- gsub(",", "_", toString(list_nested))
    # "unite" makes copy, therefore new dataset must be specified
df_all <- unite(
        data = df, lmm_group,
        list_nested, sep = "-", remove = FALSE
    )
View(df_all)
length(df_all$Sex)
```

## Choose order of factor for linear mixed model
- First element of factor level will be used as reference model
```{r}
df_all$sphere[df_all$sphere == "1.25"] <- 'r=1.25'
df_all$sphere[df_all$sphere == "2.5"] <- 'r=2.5'
df_all$sphere <- factor(df_all$sphere, levels = c("r=1.25", 'r=2.5'))
df_all$Area <- factor(df_all$Area, levels = c("IFG", "M1"))
df_all$Montage <- factor(df_all$Montage, levels = c("conv", "focal"))
df_all$Session <- factor(df_all$Session, levels = c("1", "2"))
df_all$Treatment <- factor(df_all$Treatment, levels = c("Stim", "Sham"))
df_all$Sex[df_all$Sex == "männlich"] <- 'male'
df_all$Sex[df_all$Sex == "weiblich"] <- 'female'
df_all$Sex <- factor(df_all$Sex, levels = c("female", "male"))
```

## Change level names
```{r}
names(df_all)[names(df_all) == 'coordinate space'] <- 'coordinatespace'
df_all$coordinatespace[df_all$coordinatespace == "mni2sub"] <- 'planned'
df_all$coordinatespace[df_all$coordinatespace == "sub"] <- 'actual'
df_all$coordinatespace <- factor(df_all$coordinatespace, levels = c("planned", "actual"))
names(df_all)[names(df_all) == "mean nE-field > 0 V/m"] <- 'meanNormalE'
names(df_all)
View(df_all)
```
```{r}
View(df_all)
```
# 2. Generate Models

## Generate LMM, montage in front

```{r}
# Check https://cran.r-project.org/web/packages/lme4/vignettes/lmer.pdf for model construction
# Simplest model with all fixed effects and covariates
model_Mo_old <- lmer(meanNormalE ~ 1 + Sex + Session + Treatment + Montage * sphere * coordinatespace * Area + (1 | lmm_group), data = df_all)
summary(model_Mo)

# Correlated random intercept and slope for factor sphere
model_coord_rand <- lmer(meanNormalE ~ Sex + Session + Treatment + Montage * sphere * coordinatespace * Area + (coordinatespace | lmm_group), data = df_all)
coef(summary(model_coord_rand))

# Correlated random intercept and slope for  factor sphere and electrode position
model_Mo <- lmer(meanNormalE ~ Sex + Session + Treatment + Montage * sphere * coordinatespace * Area + (sphere + coordinatespace | lmm_group), data = df_all)
coef(summary(model_Mo))

# Uncorrelated random intercept and slope for  factor sphere and electrode position
model_Mo_uncorr <- lmer(meanNormalE ~ Sex + Session + Treatment + Montage * sphere * coordinatespace * Area + (sphere + coordinatespace || lmm_group), data = df_all)
coef(summary(model_Mo_uncorr))

# Correlated random intercept and slope for  factor sphere and electrode position and Montage
model_coord_Mo_new <- lmer(meanNormalE ~ Sex + Session + Treatment + Montage * sphere * coordinatespace * Area + (sphere + coordinatespace + Montage| lmm_group), data = df_all)
coef(summary(model_coord_Mo_new))
```
# show which factors are set to 0 and which are 1 to interpret the model
```{r}
contrasts(factor(df_all$sphere))
```

## Compare models

```{r}
anova(model_Mo_old,
model_Mo,
model_coord_Mo_new,
model_Mo_uncorr,
test = "LRT")
```

## Generate Table 6

```{r}
sjPlot::tab_model(model_Mo, p.val = "kr", show.df = TRUE)
```

# generate R² values also shown in Table 6
```{r}
r2beta(model = model_Mo, partial = TRUE, method = "kr")
```
```{r}
# Model check
mo <- performance::check_model(model_Mo, panel = FALSE)
plot(mo)
```

# 3. Contrasts and Estimated Marginal Means (EMMs)

## Main effects

### Example plot (not used in paper)
```{r}
plot_df_Mo <- emmip(model_Mo, ~ Montage, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df_Mo, linearg = list(linetype = "solid"), lwd = 0,
                    dotarg = list( size = 7)) +
                    theme_bw() + 
                    ylab(ylab_LMM_text) +
                    xlab(element_blank()) +
theme(text = element_text(family = "DejaVuSans", face = "italic", size = 22)
        ) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## EMMs main effects

```{r}
emm_s.Montage_r <- emms_generate(model_Mo, path_images, "Montage")
View(emm_s.Montage_r)

emm_s.Area_r <- emms_generate(model_Mo, path_images, "Area")
View(emm_s.Area_r)

emm_s.Electrodeposition_r <- emms_generate(model_Mo, path_images, "coordinatespace")
View(emm_s.Electrodeposition_r)

emm_s.Sphere_r <- emms_generate(model_Mo, path_images, "sphere")
View(emm_s.Sphere_r)
```

## Simple effects
### Example: plot Montage model interaction over sphere
- change factors if needed
```{r}
plot_df_Mo <- emmip(model_Mo, ~ Montage | sphere, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df_Mo, linearg = list(linetype = "solid"), lwd = 0,
                    dotarg = list( size = 7)) +
                    theme_bw() + 
                    ylab(ylab_LMM_text) +
                    xlab(element_blank()) +
theme(text = element_text(family = "DejaVuSans", face = "italic", size = 22)
        ) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Generate EMMs Estimaged Marginal Means for simple effects
```{r}
emm_s.Montage_Area_r <- emms_generate(model_Mo, path_images, "Montage", "Area")
View(emm_s.Montage_Area_r)

emm_s.Montage_Sphere_r <- emms_generate(model_Mo, path_images, "Montage", "sphere")
View(emm_s.Montage_Sphere_r)

emm_s.Montage_Electrodeposition_r <- emms_generate(model_Mo, path_images, "Montage", "coordinatespace")
View(emm_s.Montage_Electrodeposition_r)

emm_s.Area_Montage_r <- emms_generate(model_Mo, path_images, "Area", "Montage")
View(emm_s.Area_Montage_r)

emm_s.Area_Electrodeposition_r <- emms_generate(model_Mo, path_images, "Area", "coordinatespace")
View(emm_s.Area_Electrodeposition_r)

emm_s.Area_Sphere_r <- emms_generate(model_Mo, path_images, "Area", "sphere")
View(emm_s.Area_Sphere_r)

emm_s.Sphere_Montage_r <- emms_generate(model_Mo, path_images, "sphere", "Montage")
View(emm_s.Sphere_Montage_r)

emm_s.Sphere_Area_r <- emms_generate(model_Mo, path_images, "sphere", "Area")
View(emm_s.Sphere_Area_r)

emm_s.Sphere_Electrodeposition_r <- emms_generate(model_Mo, path_images, "sphere", "coordinatespace")
View(emm_s.Sphere_Electrodeposition_r)

emm_s.Electrodeposition_Montage_r <- emms_generate(model_Mo, path_images, "coordinatespace", "Montage")
View(emm_s.Electrodeposition_Montage_r)

emm_s.Electrodeposition_Area_r <- emms_generate(model_Mo, path_images, "coordinatespace", "Area")
View(emm_s.Electrodeposition_Area_r)

emm_s.Electrodeposition_Sphere_r <- emms_generate(model_Mo, path_images, "coordinatespace", "sphere")
View(emm_s.Electrodeposition_Sphere_r)
```


## SIMPLE CONTRASTS

### plot Montage model interaction over sphere and electrodes
- Example for automatic plotting of interaction
```{r}
plot_df_Mo <- emmip(model_Mo, ~ Montage | coordinatespace * sphere, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df_Mo, linearg = list(linetype = "solid"), lwd = 0, dotarg = list( size = 3)) +
  theme_bw() + 
  ylab(ylab_LMM_text) +
  xlab(element_blank()) +
  ylim(-0.01, 0.16) +
  theme(text = element_text(family = "DejaVuSans", size = 24)) +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
View(plot_df_Mo)
```

### Generate EMMs Estimaged Marginal Means for simple contrasts
```{r}
emm_s.Montage_Electrodeposition_Sphere_r <- emms_generate(model_Mo, path_images, "Montage", "coordinatespace", "sphere")
View(emm_s.Montage_Electrodeposition_Sphere_r)

emm_s.Electrodeposition_Montage_Area_r <- emms_generate(model_Mo, path_images, "coordinatespace", "Montage", "Area")
View(emm_s.Electrodeposition_Montage_Area_r)

emm_s.Montage_Sphere_Electrodeposition_r <- emms_generate(model_Mo, path_images,"Montage", "sphere", "coordinatespace")
View(emm_s.Montage_Sphere_Electrodeposition_r)

emm_s.Sphere_Montage_Electrodeposition_r <- emms_generate(model_Mo, path_images, "sphere", "Montage", "coordinatespace")
View(emm_s.Sphere_Montage_Electrodeposition_r)

emm_s.Electrodeposition_Montage_Sphere_r <- emms_generate(model_Mo, path_images, "coordinatespace","Montage", "sphere")
View(emm_s.Electrodeposition_Montage_Sphere_r)
```


## Simple contras all factors
### Example: plot Montage model interaction over sphere and electrodes
- Automatic example for simple contrast of all factors
- Change first factor to change contrast (Example: Montage -> Area)
```{r}
plot_df_Mo <- emmip(model_Mo, ~ Montage | coordinatespace * sphere * Area, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df_Mo, linearg = list(linetype = "solid"), lwd = 0, dotarg = list(size = 7)) +
  theme_bw() +
  ylab("95% CI EMM Measurement V/m") +
  xlab(element_blank()) +
  ylim(0.00, 0.20) +
  theme(text = element_text(family = "DejaVuSans", face = "italic", size = 22)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Generate EMMs Estimaged Marginal Means for simple contrasts (all factors)
```{r}
emm_s.Montage_Electrodeposition_Sphere_Area_r <- emms_generate(model_Mo, path_images, "Montage", "coordinatespace", "sphere", "Area")
View(emm_s.Montage_Electrodeposition_Sphere_Area_r)

emm_s.Area_Montage_Electrodeposition_Sphere_r  <- emms_generate(model_Mo, path_images, "Area", "Montage", "coordinatespace", "sphere")
View(emm_s.Area_Montage_Electrodeposition_Sphere_r)

emm_s.Sphere_Montage_Electrodeposition_Area_r  <- emms_generate(model_Mo, path_images, "sphere", "Montage", "coordinatespace", "Area")
View(emm_s.Sphere_Montage_Electrodeposition_Area_r)

emm_s.Electrodeposition_Montage_Sphere_Area_r <- emms_generate(model_Mo, path_images, "coordinatespace", "Montage", "sphere", "Area")
View(emm_s.Electrodeposition_Montage_Sphere_Area_r)
```

# 4. Plot Contrast and EMMs
## Main factors
### Plot Main Factor Montage
```{r}
plot_posthoc(df_all, emm_s.Montage_r, Montage, "blue", "Montage")
ggsave(file.path(path_images,"R_EMM_Montage.png"), width = width*faktor, height = height*faktor, dpi = dpi)
```

### Plot Main Factor Area
```{r}
plot_posthoc(df_all, emm_s.Area_r, Area, "blue", "Area")
ggsave(file.path(path_images,"R_EMM_Area.png"), width = width*faktor, height = height*faktor, dpi = dpi)
```

### Plot Main Factor Sphere
```{r}
plot_posthoc(df_all, emm_s.Sphere_r, sphere, "blue", "ROI radius in cm")
ggsave(file.path(path_images,"R_EMM_Sphere.png"), width = width*faktor, height = height*faktor, dpi = dpi)
```

### Plot Main Factor Electrode position
```{r}
plot_posthoc(df_all, emm_s.Electrodeposition_r, coordinatespace, "blue","Electrode position")
ggsave(file.path(path_images,"R_EMM_Electrode.png"), width = width*faktor, height = height*faktor, dpi = dpi)
```

## Plot simple contrasts

### Contrast Electrode position with interaction of Montage and ROI radius
```{r}
g1 <- plot_posthoc_multi(df_all, emm_s.Electrodeposition_Montage_Sphere_r, coordinatespace,
"Montage", "conv", "sphere", "r=1.25",
"1", "Electrode position", Montage, sphere, "blue")
g2 <- plot_posthoc_multi(df_all, emm_s.Electrodeposition_Montage_Sphere_r, coordinatespace,
"Montage", "conv", "sphere", "r=2.5",
"2", "Electrode position", Montage, sphere,"blue")
g3 <- plot_posthoc_multi(df_all, emm_s.Electrodeposition_Montage_Sphere_r, coordinatespace,
"Montage", "focal", "sphere", "r=1.25",
"3", "Electrode position", Montage, sphere,"blue")
g4 <- plot_posthoc_multi(df_all, emm_s.Electrodeposition_Montage_Sphere_r, coordinatespace,
"Montage", "focal", "sphere", "r=2.5",
"4", "Electrode position", Montage, sphere,"blue")

grid_arrange(g1,g2,g3,g4,"Electrode position","R_EMM_Electrodes_Montage_Sphere_color.png")
```

## Contrast ROI radius with interaction of Montage and Electrode position
```{r}
b1 <- plot_posthoc_multi(df_all, emm_s.Sphere_Montage_Electrodeposition_r, sphere,
"Montage", "conv", "coordinatespace", "planned",
"1", "ROI radius", Montage, coordinatespace,"blue")
b2 <- plot_posthoc_multi(df_all, emm_s.Sphere_Montage_Electrodeposition_r, sphere,
"Montage", "conv", "coordinatespace", "actual",
"2", "ROI radius", Montage, coordinatespace,"blue")
b3 <- plot_posthoc_multi(df_all, emm_s.Sphere_Montage_Electrodeposition_r, sphere,
"Montage", "focal", "coordinatespace", "planned",
"3", "ROI radius", Montage, coordinatespace,"blue")
b4 <- plot_posthoc_multi(df_all, emm_s.Sphere_Montage_Electrodeposition_r, sphere,
"Montage", "focal", "coordinatespace", "actual",
"4", "ROI radius", Montage, coordinatespace,"blue")

grid_arrange(b1, b2, b3, b4, "ROI radius in cm","R_EMM_Sphere_Montage_Electrodes_color.png")

```

## Contrast Montage with interaction of ROI radius and Electrode position
```{r}
c1 <- plot_posthoc_multi(df_all, emm_s.Montage_Sphere_Electrodeposition_r, Montage,
"sphere", "r=1.25", "coordinatespace", "planned",
"1", "Montage", coordinatespace, sphere,"blue")
c2 <- plot_posthoc_multi(df_all, emm_s.Montage_Sphere_Electrodeposition_r, Montage,
"sphere", "r=2.5", "coordinatespace", "planned",
"2", "Montage", coordinatespace, sphere,"blue")
c3 <- plot_posthoc_multi(df_all, emm_s.Montage_Sphere_Electrodeposition_r, Montage,
"sphere", "r=1.25", "coordinatespace", "actual",
"3", "Montage", coordinatespace, sphere,"blue")
c4 <- plot_posthoc_multi(df_all, emm_s.Montage_Sphere_Electrodeposition_r, Montage,
"sphere", "r=2.5", "coordinatespace", "actual",
"4", "Montage", coordinatespace, sphere,"blue")

c1 + c2 + c3 + c4
grid_arrange(c1, c2, c3, c4, "Montage", "R_EMM_Montage_Electrodes_Sphere_color.png")
```


