---
title: "R script for coordinate differences "
author: Filip Niemann
date: 20.09.2023
output: html_document
---

# load librarys
```{r}
library(lattice)
library(ggplot2)
library(arrow)
library(dplyr)
library(rstatix)
library(emmeans)
library(knitr)
library(lme4)
library(afex)
library(tidyr)
library(sjmisc)
library(lmerTest)
library(httpgd)
library(r2glmm)
library(see)
library(patchwork)
library(gridExtra)
library(cowplot)
library(tidyverse)
library(grid)
library(here)

# If you use visual code, images will be shown in extra browser window
hgd()
hgd_browse()
```

```{r, echo=FALSE}
library(rstudioapi)
source_path <- dirname(getSourceEditorContext()$path)
# Should be the path on your local machine
View(source_path)
# Use given functions to calculate estimated marginal means and plot them
source(file.path(source_path, "plot_emmip.R"))
# Path to figure and table files (some are given, some will be created)
path_source_data <- (file.path(source_path, "Source_Data"))
```


# Variables for tables

```{r}
# Style and path variables
digits <- 2
digits2 <- 4
path_images <- file.path(source_path, "Images_and_Tables")
width <- 10
height <- 7
dpi <- 150
fontsize <- 24
ylab_text <- "Difference (|mm|) \n planned - actual"
ylab_LMM_text <- "95% CI EMM (|mm|)"
```

```{r, echo=FALSE}
# Table for all electrode coordinate values
data <- file.path(path_source_data, "df_all_diff_abs_Coordinates_long_format.feather")
df_new <- arrow::read_feather(data)
View(df_new)
```

```{r}
# Delete additional, not used rows in data
df_new <- df_new[!grepl("konv-tDCS IFG letters", df_new$Exp), ]
``` 


## Change Montage naming

```{r}
df_new <- within(df_new, rm(Montage))
df_new <- df_new %>% mutate(Montage = case_when(
  grepl(pattern = "konv", x = Exp) ~ "conv",
  grepl(pattern = "focal", x = Exp) ~ "focal")
)
```

## Load  df with Session and control same length

```{r, echo=FALSE}
df_new %>% group_by(Exp, Treatment, Electrode, `coordinate space`) %>% summarise(n())
```


# Choose factor, first factor values dummy coded as 0

```{r}
df_new$Electrode <- factor(df_new$Electrode, levels = c("y", "z", "x"))
df_new$Area <- factor(df_new$Area, levels = c("IFG", "M1"))
df_new$Montage[df_new$Montage == "conv"] <- 'conv'
df_new$Montage[df_new$Montage == "focal"] <- 'focal'
df_new$Montage <- factor(df_new$Montage, levels = c("conv", "focal"))
df_new$Treatment <- factor(df_new$Treatment, levels = c("Stim", "Sham"))
df_new$Sex[df_new$Sex == "männlich"] <- 'male'
df_new$Sex[df_new$Sex == "weiblich"] <- 'female'
df_new$Sex <- factor(df_new$Sex, levels = c("female", "male"))
df_new$Session <- factor(df_new$Session, levels = c("1", "2"))
View(df_new)
```

## Create index for random variable (are similar named in each experiment)

```{r}
list_nested <- c("Exp", "index")
lmm_group <- gsub(",", "_", toString(list_nested))
# "unite" makes copy, therefore new dataset must be specified
df_new <- unite(
  data = df_new, lmm_group,
  list_nested, sep = "-", remove = FALSE
)
```

# 2. Generate Models

## Generate LMM, montage in front


```{r}
# Model without interaction effects between electrode position and experimental montages
lmer_model <- lmer(diff ~ 1 + Sex + Treatment + Session + Electrode + Montage * Area + (1 | lmm_group), data = df_new)
#summary(lmer_model)

#Model with interaction effects between electrode position and experimental montages
lmer_model_Area <- lmer(diff ~ 1 + Sex + Treatment + Session + Area * Montage * Electrode + (1 | lmm_group), data = df_new)
#summary(lmer_model_Area)
```

## Compare models

```{r}
# Choose winning model with highest AIC and BIC score
anova(lmer_model, lmer_model_Area, test = "LRT")
```

```{r}
sjPlot::tab_model(lmer_model_Area,, show.ci = 0.95, show.p = TRUE, show.df = TRUE)
```

# Plot simple effects of the model
```{r}
sjPlot::plot_model(lmer_model_Area,
                   show.values = TRUE,
                   show.p = TRUE,
                   title = "Electrode coordinates")
```
## Generate Table 2
```{r}
sjPlot::tab_model(lmer_model_Area, p.val = "kr", show.df = TRUE)
```

# generate R² values also shown in Table 2

```{r}
r2beta(model = lmer_model_Area, partial = TRUE, method = "nsj")
```


# 3. Contrasts and Estimated Marginal Means (EMMs)

## Main effects

### Example plot (not used in paper)
```{r}
plot_df <- emmip(lmer_model_Area, ~ Montage, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df, linearg = list(linetype = "solid"), lwd = 0, dotarg = list( size = 7)) +
  theme_bw() + 
  ylab(ylab_LMM_text) +
  xlab(element_blank()) +
  theme(text = element_text(family = "DejaVuSans", face = "italic", size = 22)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## EMMs main effects

```{r}
emm_s.Area_r <- emms_generate(lmer_model_Area, path_images, "Area")
View(emm_s.Area_r)

emm_s.Montage_r <- emms_generate(lmer_model_Area, path_images, "Montage")
View(emm_s.Montage_r)

emm_s.Electrode_r <- emms_generate(lmer_model_Area, path_images, "Electrode")
View(emm_s.Electrode_r)
```

## Simple effects
### Example: plot Montage model interaction over Area
- change factors if needed
```{r}
plot_df <- emmip(lmer_model_Area, ~ Montage | Area, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df, linearg = list(linetype = "solid"), lwd = 0,
             dotarg = list( size = 7)) +
             theme_bw() +
             ylab(ylab_LMM_text) +
             xlab(element_blank()) +
theme(text = element_text(family = "DejaVuSans", face = "italic", size = 22)) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Generate EMMs for simple effects
```{r}
emm_s.Montage_Electrode_r <- emms_generate(lmer_model_Area, path_images, "Montage", "Electrode")
View(emm_s.Montage_Electrode_r)

emm_s.Montage_Area_r <- emms_generate(lmer_model_Area, path_images, "Montage", "Area")
View(emm_s.Montage_Area_r)

emm_s.Area_Electrode_r <- emms_generate(lmer_model_Area, path_images, "Area", "Electrode")
View(emm_s.Area_Electrode_r)

emm_s.Area_Montage_r <- emms_generate(lmer_model_Area, path_images, "Area", "Montage")
View(emm_s.Area_Montage_r)

emm_s.Electrode_Area_r <- emms_generate(lmer_model_Area, path_images, "Electrode", "Area")
View(emm_s.Electrode_Area_r)

emm_s.Electrode_Montage_r <- emms_generate(lmer_model_Area, path_images, "Electrode", "Montage")
View(emm_s.Electrode_Montage_r)
```

## SIMPLE CONTRASTS

### plot Montage model interaction over sphere and electrodes
- example for automatic plotting of interaction
```{r}
plot_df <- emmip(lmer_model_Area, ~ Area | Montage * Electrode, plotit = FALSE, CIs = TRUE)
emmip_ggplot(plot_df, linearg = list(linetype = "solid"), lwd = 0, dotarg = list( size = 3)) +
  theme_bw() + 
  ylab(ylab_LMM_text) +
  xlab(element_blank()) +
  ylim(0.00, 16) +
  theme(text = element_text(family = "DejaVuSans", size = 24)) +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))
View(plot_df)
```

### Generate EMMs Estimaged Marginal Means for simple contrasts

```{r}
emm_s.Montage_Electrode_Area_r <- emms_generate(lmer_model_Area, path_images, "Montage", "Electrode", "Area")
View(emm_s.Montage_Electrode_Area_r)

emm_s.Montage_Area_Electrode_r <- emms_generate(lmer_model_Area, path_images, "Montage", "Area", "Electrode")
View(emm_s.Montage_Area_Electrode_r)

emm_s.Area_Electrode_Montage_r <- emms_generate(lmer_model_Area, path_images, "Area", "Electrode", "Montage")
View(emm_s.Area_Electrode_Montage_r)

emm_s.Electrode_Montage_Area_r <- emms_generate(lmer_model_Area, path_images, "Electrode", "Montage","Area")
View(emm_s.Electrode_Montage_Area_r)

emm_s.Electrode_Area_Montage_r <- emms_generate(lmer_model_Area, path_images, "Electrode", "Area","Montage")
View(emm_s.Electrode_Area_Montage_r)
```

# 4. Plot Contrast and EMMs
## Main factors
### Plot Main Factor Montage

```{r}
plot_posthoc(df_new, emm_s.Montage_r, Montage, "blue")
ggsave(file.path(path_images,"Interaction_plot_Montage_Electrodes.png"), width = width, height = height, dpi = dpi)

```

### Plot Main Factor Area

```{r}
plot_posthoc(df_new, emm_s.Area_r, Area, "blue")
ggsave(file.path(path_images,"Interaction_plot_Area_Electrodes.png"), width = width, height = height, dpi = dpi)
```

## Plot simple contrasts

### Contrast Montage with interaction of Area and electrode difference

```{r}
g1 <- plot_posthoc_multi(df_new, emm_s.Montage_Electrode_Area_r, Montage,
"Area", "IFG", "Electrode", "x",
"1", "Area", Area, Electrode, "blue")

g2 <- plot_posthoc_multi(df_new, emm_s.Montage_Electrode_Area_r, Montage,
"Area", "IFG", "Electrode", "y",
"2", "Area", Area, Electrode, "blue")

g3 <- plot_posthoc_multi(df_new, emm_s.Montage_Electrode_Area_r, Montage,
"Area", "IFG", "Electrode", "z",
"3", "Area", Area, Electrode, "blue")

g4 <- plot_posthoc_multi(df_new, emm_s.Montage_Electrode_Area_r, Montage,
"Area", "M1", "Electrode", "x",
"4", "", Area, Electrode, "blue")

g5 <- plot_posthoc_multi(df_new, emm_s.Montage_Electrode_Area_r, Montage,
"Area", "M1", "Electrode", "y",
"5", "Montage", Area, Electrode, "blue")

g6 <- plot_posthoc_multi(df_new, emm_s.Montage_Electrode_Area_r, Montage,
"Area", "M1", "Electrode", "z",
"6", "", Area, Electrode, "blue")

# Combine all simple plots
g1 + g2 + g3 + g4 + g5 + g6
ggsave(file.path(path_images,"Interaction_plot_Montage_Electrode_Area.png"), width = height*2, height = width,, dpi = dpi)

```

### Contrast Area with interaction of Montage and electrode difference

```{r}
b1 <- plot_posthoc_multi(df_new, emm_s.Area_Electrode_Montage_r, Area,
"Montage", "conv", "Electrode", "x",
"1", "Area", Montage, Electrode, "blue")

b2 <- plot_posthoc_multi(df_new, emm_s.Area_Electrode_Montage_r, Area,
"Montage", "conv", "Electrode", "y",
"2", "Area", Montage, Electrode, "blue")

b3 <- plot_posthoc_multi(df_new, emm_s.Area_Electrode_Montage_r, Area,
"Montage", "conv", "Electrode", "z",
"3", "Area", Montage, Electrode, "blue")

b4 <- plot_posthoc_multi(df_new, emm_s.Area_Electrode_Montage_r, Area,
"Montage", "focal", "Electrode", "x",
"4", "", Montage, Electrode, "blue")

b5 <- plot_posthoc_multi(df_new, emm_s.Area_Electrode_Montage_r, Area,
"Montage", "focal", "Electrode", "y",
"5", "Area", Montage, Electrode, "blue")

b6 <- plot_posthoc_multi(df_new, emm_s.Area_Electrode_Montage_r, Area,
"Montage", "focal", "Electrode", "z",
"6", "", Montage, Electrode, "blue")

# Combine all simple plots
b1 + b2 + b3 + b4 + b5 + b6
ggsave(file.path(path_images,"Interaction_plot_Area_Electrode_Montage.png"), width = height*2, height = width, dpi = dpi)
```

